Knowledge Representation for Service Robots: Supply and Demand

### 2012-10-08 krs-survey.tex
Integration of Moritz's comments

    Central to this effort is to develop an understanding of the relationship
    between the knowledge, the perception, and the action of [\ldots] a robot. The
    sorts of questions we want to be able to answer are

@@

    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.

@@

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language, semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the thesis, give a second look to
this distinction.

@@

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question. In this thesis,
we will associate data to low-level material like raw sensor output, and
information to uncontextualised symbolic facts.

To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, book)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

@@

The aim of this imaginary scenario (that has not been implemented, neither in
simulation nor on a robot) is to materialise early in this paper the context
and challenges of knowledge representation and manipulation for service, social
and interactive robotics. It underlines the place, the role and the need of
knowledge in a near-future, everyday situation where several robots and humans
co-exist and cooperate. This scenario will also be a source of support examples
for later sections of the thesis. Also, we have emphasised several keywords in
this description: we will come back to them in the next section to
explain them formally and relate them to each other.

@@

A ``first-order'' analysis of this task leads to a rough partition of the
required \emph{representation} abilities:

@@

These last assertions imply several other capabilities: ``cooking often takes
place in the kitchen'' implies that representation of both uncertainty and
likelihood is desirable. The fact that cooking is associated to a place further
implies that the system models locations and is able to attach \emph{thematic
relations} to concepts (here, the likely location of the cooking action).

@@

Cooperation in our scenario can intervene at many places. For instance, an
agent may want to inform another one about the number of eggs that are
necessary for the brownie. This \emph{helping} behaviour makes sense only if
the first agent knows that the recipient agent both needs the information but
does not know it. This in turn requires the robot to be able to model the
knowledge of the other agents: to think \emph{from the perspective} of another
agent (an idea that is related to the availability of a theory of mind, we will
come back to it later on).

@@

It can be noted that the relation between expressiveness and reasoning
complexity is fragile: for instance, adding the following axiom \stmt{farFrom
disjointProperty near} (that states that two individuals can not be at the same
time near and far from each other) changes the expressiveness power of the ORO
Common-Sense ontology (presented at chapter~\ref{chapt|oroserver}) from
$\mathcal{SHOIQ(D)}$ to $\mathcal{SROIQ(D)}$\footnote{See
appendix~\ref{chapt|dl} for a brief explanation of this notation}: this
seemingly innocuous assertion change the complexity class of the whole
ontology, and the concept satisfiability reasoning problem switches from a {\it
NExpTime-complete} problem to a {\it NExpTime-hard} problem (\ie, \emph{at
least} as hard as the hardest problem in {\it NExpTime}).

This ``instability'' has practical consequences on run-time performances on
the robot because a light alteration of the knowledge structure can lead to
very noticeable performance drops.

@@

Symbolic representation of entities placement can be absolute or relative. The
relation \concept{isOn}, for example, leads to absolute statements: the
validity of the relation is independent of the nature of its subjects and
objects, and is also independent of the observer.

On the contrary, the relation \concept{nextTo} is relative, and depends on the
relative size of the subject and the object. Two houses distant of 2 meters
from each other can be considered as next to each other. Two ants separated by
2 meters are not next to each other.

The relation \concept{leftOf} is another example of relative spatial relation,
this time because it depends on the observer viewpoint.

@@

We see context representation as one of the main challenge of knowledge
representatio in general, and we will further discuss the importance and issues
brought by context modeling in the conclusion of the thesis.

@@

One downside is that the knowledge base never contains explicitly the complete
set of beliefs of the robot. This limits for instance the ability for the robot
to react to logical conditions that involve facts that are lazily evaluated
(``trigger a callback when object A is behind object B'' would not be triggered
with a purely lazy evaluation strategy for spatial relations).

@@

\paragraph{Projection task}: determining whether or not some condition while
hold after a sequence of actions. The projection task is a typical
non-monotonic reasoning task, since at each step, the system must add but also
retract beliefs, as defined in the tasks post-conditions.

@@

Being actually implemented on real service robots, all the symbolic knowledge
representation systems that we review in this study have some kind of grounding
process. Numerous approaches exist, like amodal
\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
representations~\cite{Alami2011, Mavridis2006}, semantic maps
(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planing and object classification~\cite{Lorken2008,
Varadarajan2011}.

@@

@@

\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{1.5cm}}
\toprule
{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
\midrule
ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & \cite{Jacobsson2008} \\
GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & & \cite{Mavridis2006} \\
Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & \cite{Sabri2011} \\
%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
%ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
%Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
% (based on {\sc KnowRob} & & (JSK) \\

@@

The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
representation language with natural language processing, task planing and
motion planing.

Knowledge representation relies on \emph{Action Language C}, itself based on
\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
are syntactically close to Prolog, are based on \emph{stable models} of logic
programs, and support non-monotonic reasoning. Default and non-monotonic
reasoning has been especially researched within the Ke Jia project for symbolic
task planing~\cite{Ji2011} and underspecified natural language processing.

@@

\begin{landscape}
\begin{table}\tiny
\begin{center}
\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}|p{2cm}}
\toprule
\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}       & ORO \cite{Lemaignan2010}                      \\

@@

\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL + Horn clauses             & {\sc CycL}                    & DL (OWL)                                      \\
                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 &                               &                               & OWA                                           \\
                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         & + (\emph{candidate} entities) &                               &                                               \\
                                                                    & Meta-cognition &                           &                             &                             &                             & ++                                 & + (transformation rules)                        &                               &                               & ++ (reification, taxonomy walking)            \\

@@

\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++                            &                               & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})\\
                                                               & Time representation &                           &                             & ++ (snapshots)              &                             &                                    & ++ (event oriented)                             & +                             &                               &                                               \\
                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            & +++                           &                               &                                               \\
                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          & ++                            & ++ (microtheories)            &                                               \\
                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 &                               &                               & ++ (Theory of Mind \cite{Warnier2012a})       \\

@@

                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 &                               & +                             & +                                             \\

@@

\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & +                             & ++                            & ++                                            \\
                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 &                               &                               & ++ (TBox alteration)                          \\

@@

                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           &                                    &                                                 &                               &                               &                                               \\

@@

                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             &                                    & ++                                              &                               &                               & + (explanation)                               \\
                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +                             & {\it++}                       & +++ ({\em HATP})                              \\

@@

\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 &                               & ++ ({\em ambient intelligence})& ++                                           \\
                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 &                               & + (template-based)            & +++ ({\em Dialogs} \cite{Lemaignan2011a})     \\

@@

                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & +++ (bottom-up)               & +++ \cite{Loutfi2008}         & ++ ({\em amodal model} \cite{Lemaignan2012})  \\

@@

\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          & +                             & ++ (tuple space)              &                                               \\
                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++                            & ++ (tuple space)              & ++ (semantic events)                          \\
                                                          & Monitoring and debugging &                           &                             &                             &                             &                                    &                                                 &                               &                               & + ({\em remote visualisation})                \\

@@

The interleaving of discreet symbolic models with continuous geometric models
remains a major challenge, and techniques vary a lot: besides the CAST proxies
and the GSM amodal model, most systems directly extract physical attributes of
the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
adopt a top-down approach where local geometric models are set up on-demand to
compute symbolic properties like relative locations.

In our approach, ORO has tight (but unidirectional) links with SPARK, the
component in charge of geometric modeling and reasoning. Similarly to the GSM
approach, SPARK is an amodal model of the environment we can manipulate (3D
motion planing, for instance takes place in MHP, the twin brother of SPARK).

@@

Several approaches for building contextualized knowledge have been presented in
this thesis: symbolic environment interpretation, perspective taking and
independent mental state for each agent, grounded natural language resolution,
self-awareness of its own activity. Much remains to be done for a robot to
actually identify its current context as well as contexts that may be referred
to.

We will further develop this important topic in the conclusion of the thesis.

@@

The two main tasks that require time reasoning are task planing and sequence
recognition (in particular, action recognition). Systems that directly
integrate task planing (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
fluents in {\sc KnowRob}) to represent time.

@@

We have already explained how perspective awareness enables advanced cognitive
capabilities like a theory of mind, and this is probably the main contribution
of this thesis regarding techniques for knowledge representation.

>>>>

    ``Central to this effort is to develop an understanding of the relationship
    between the knowledge, the perception, and the action of [\ldots] a robot. The
    sorts of questions we want to be able to answer are

@@

    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.''

@@

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language, semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the article, give a second look to
this distinction.

@@

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question. In this article,
we will associate data to low-level material like raw sensor output, and
information to uncontextualised symbolic facts.

To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, coords)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

@@

The aim of this imaginary scenario (that has not been implemented, neither in
simulation nor on a robot) is to materialise early in this paper the context
and challenges of knowledge representation and manipulation for service, social
and interactive robotics. It underlines the place, the role and the need of
knowledge in a near-future, everyday situation where several robots and humans
co-exist and cooperate. This scenario will also be a source of support examples
for later sections of the article. Also, we have emphasised several keywords in
this description: we will come back to them in the next section to
explain them formally and relate them to each other.

@@

A \textit{first order} analysis of this task leads to a rough partition of the
required \emph{representation} abilities:

@@

These last assertions imply several other capabilities: ``cooking \emph{often} takes
place in the kitchen'' implies that representation of both uncertainty and
likelihood is desirable. The fact that cooking is associated to a place further
implies that the system models locations and is able to attach \emph{thematic
relations} to concepts (here, the likely location of the cooking action).

@@

Cooperation in our scenario can intervene at many places. For instance, an
agent may want to inform another one about the number of eggs that are
necessary for the brownie. This \emph{helping} behaviour makes sense only if
the first agent knows that the recipient agent both needs the information but
does not know it. This in turn requires the robot to be able to access its own
knowledge (\emph{meta-cognition}), as well as to model the knowledge of the
other agents (to think \emph{from the perspective} of another agent, an idea
that is related to the availability of a theory of mind, we will come back to
it later on).

@@

It can be noted that the relation between expressiveness and reasoning
complexity is fragile: for instance, adding the following axiom \stmt{farFrom
disjointProperty near} that states that two individuals can not be at the same
time near and far from each other, while looking semantically seemingly
innocuous, actually introduces what is called a \emph{complex role} that leads
to a much harder reasoning problem: concept satisfiability, for instance,
switches from a {\it NExpTime-complete} problem to a {\it NExpTime-hard}
problem (\ie, \emph{at least} as hard as the hardest problem in {\it
NExpTime}).

This kind of behaviours have practical consequences on run-time performances on
the robot, and require careful design of the knowledge structure to be mitigated.

@@

Symbolic representations of entities placement are either topological or
directional. The topological relation \concept{isOn}, for example, leads to
absolute statements: the validity of the relation is independent of the nature
of its subjects and objects, and is also independent of the observer. These
relations are also sometimes qualified as \emph{allocentric}.

On the contrary, the directional relation \concept{nextTo} is relative, and
depends on the relative size of the subject and the object. Two houses distant
of 2 meters from each other can be considered as next to each other. Two ants
separated by 2 meters are not next to each other.

The relation \concept{leftOf} is another example of relative spatial relation,
this time because it depends on the observer viewpoint. These relations are
also called \emph{egocentric}.

@@

We see context representation as one of the main challenge of knowledge
representatio in general, and we will further discuss the importance and issues
brought by context modeling in the conclusion of the paper.

@@

One downside is that the knowledge base never contains explicitly the complete
set of beliefs of the robot. This limits for instance the ability for the robot
to react to logical conditions that involve facts that are lazily evaluated:
``trigger a callback when object A is behind object B'' would not be triggered
with a purely lazy evaluation strategy for spatial relations. This can be
mitigated by selecting (possibly dynamically) a subset of relations that are to
be strictly evaluated.

@@

\paragraph{Projection task}: determining whether or not some condition
hold after a sequence of actions. The projection task is a typical
non-monotonic reasoning task, since at each step, the system must add but also
retract beliefs, as defined in the tasks post-conditions.

@@

Being actually implemented on real service robots, all the symbolic knowledge
representation systems that we review in this study have some kind of grounding
process. Numerous approaches exist, like amodal
\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
representations~\cite{Alami2011, Mavridis2006}, semantic maps
(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planning and object classification~\cite{Lorken2008,
Varadarajan2011}.

@@

A last qualitative facet relates to the possible bias of an ontology
regarding the domain it pretends to cover. For instance, Wikipedia (and
consequently, DBPedia and YAGO) is known to have a bias towards contents
related to people, geography or historic events. {\sc OpenCYC} also suffer of
non-homogeneous concept coverage.

@@

\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{1.5cm}}
\toprule
{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
\midrule
ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & \cite{Jacobsson2008} \\
GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & & \cite{Mavridis2006} \\
Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & \cite{Sabri2011} \\
%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
ORO & Formal & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) + {\sc Pellet} & \cite{Lemaignan2010} \\
OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
%Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
% (based on {\sc KnowRob} & & (JSK) \\

@@

The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
representation language with natural language processing, task planning and
motion planning.

Knowledge representation relies on \emph{Action Language C}, itself based on
\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
are syntactically close to Prolog, are based on \emph{stable models} of logic
programs, and support non-monotonic reasoning. Default and non-monotonic
reasoning has been especially researched within the Ke Jia project for symbolic
task planning~\cite{Ji2011} and underspecified natural language processing.

@@

\begin{landscape}
\begin{table}\tiny
\begin{center}
\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}}
\toprule
\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & ORO \cite{Lemaignan2010}                      & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}        \\

@@

\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL (OWL)                                      & DL + Horn clauses             & {\sc CycL}                     \\
                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 & OWA                                           &                               &                                \\
                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         &                                               & + (\emph{candidate} entities) &                                \\
                                                                    & Meta-cognition &                           &                             &                             &                             & +                                  & + (transformation rules)                        & ++ (reification, taxonomy walking)            &                               &                                \\

@@

\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})& ++         &                               \\
                                                               & Time representation &                           &                             & ++ (snapshots)              &                             & + (Allen's intervals, via \emph{computables} & ++ (event oriented)                   &                                               & +                             &                                \\
                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            &                                               & +++                           &                                \\
                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          &                                               & ++                            & ++ (microtheories)             \\
                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 & ++ (Theory of Mind \cite{Warnier2012a})       &                               &                                \\

@@

                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 & +                                             &                               & +                              \\

@@

\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & ++                                            & +                             & ++                             \\
                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 & ++ (TBox alteration)                          &                               &                                \\

@@

                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           & + \cite{Tenorth2012}               &                                                 &                                               &                               &                                \\

@@

                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             & + \cite{Tenorth2012}                                   & ++                                              & + (explanation)                               &                               &                                \\
                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +++ ({\em HATP})                              & +                             & {\it++}                        \\

@@

\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 & ++                                            &                               & ++ ({\em ambient intelligence})\\
                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 & +++ ({\em Dialogs} \cite{Lemaignan2011a})     &                               & + (template-based)             \\

@@

                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & ++ ({\em amodal model} \cite{Lemaignan2012})  & +++ (bottom-up)               & +++ \cite{Loutfi2008}          \\

@@

\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          &                                               & +                             & ++ (tuple space)               \\
                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++ (semantic events)                          & ++                            & ++ (tuple space)               \\
                                                          & Monitoring and debugging &                           &                             &                             &                             & +(Prolog tracer, visualisations...)&                                                 & + ({\em remote visualisation})                &                               &                                \\

@@

The interleaving of discrete symbolic models with continuous geometric models
remains a major challenge, and techniques vary a lot: besides the CAST proxies
and the GSM amodal model, most systems directly extract physical attributes of
the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
adopt a top-down approach where local geometric models are set up on-demand to
compute symbolic properties like relative locations.

In our approach, ORO has tight (but unidirectional) links with SPARK, the
component in charge of geometric modeling and reasoning. Similarly to the GSM
approach, SPARK is an amodal model of the environment we can manipulate (3D
motion planning, for instance takes place in MHP, the twin brother of SPARK).

@@

Several approaches for building contextualized knowledge have been presented in
this article: symbolic environment interpretation, perspective taking and
independent mental state for each agent, grounded natural language resolution,
self-awareness of its own activity. Much remains to be done for a robot to
actually identify its current context as well as contexts that may be referred
to.

We will further develop this important topic in the conclusion of the paper.

@@

The two main tasks that require time reasoning are task planning and sequence
recognition (in particular, action recognition). Systems that directly
integrate task planning (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
fluents in {\sc KnowRob}) to represent time.

@@

We have already explained how perspective awareness enables advanced cognitive
capabilities like a theory of mind, and this is probably the main contribution
of this paper regarding techniques for knowledge representation.
